On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-03-27 manuscripts/PNAS/chazthm_pnas.tex
more SR up to page 2

Previous theoretical work on Problem \ref{InverseProblem} for the noiseless case $\eta =0$ (e.g., \cite{li2004analysis, Georgiev05, Aharon06, Hillar15}) has shown that a solution (when it exists) is indeed unique provided the $\mathbf{x}_i$ are sufficiently diverse and the generating matrix $\mathbf{A}$ satisfies the \textit{spark condition} from compressed sensing (CS):
\begin{align}\label{SparkCondition}
\mathbf{A}\mathbf{x}_1 = \mathbf{A}\mathbf{x}_2 \implies \mathbf{x}_1 = \mathbf{x}_2, \ \ \ \text{for all $k$-sparse } \mathbf{x}_1, \mathbf{x}_2,
\end{align}
%
which would be, in any case, a necessary condition for uniqueness if it were known only that $\mathbf{x}_i$ are $k$-sparse.

Our main technical finding is that a sufficient diversity of measurements generated from any dictionary satisfying the spark condition 
has a stable sparse representation.  More specifically, such dictionaries are uniquely identifiable from as few as \mbox{$N = m + m(k-1){m \choose k}$} sparse linear combinations of its columns up to an error that is linear in the measurement noise (Thm.~\ref{DeterministicUniquenessTheorem}). In fact, provided $(n, m, k)$ satisfy the CS inequality \eqref{CScondition}, in almost all cases the dictionary learning problem is well-posed given enough data (Cor.~\ref{ProbabilisticCor}). Importantly, these guarantees hold without assuming the recovered matrix satisfies the spark condition. Our results  also apply when only an upper bound on the number of dictionary elements is known (Thm.~\ref{DeterministicUniquenessTheorem2}). We hope the explicit, algorithm-independent criteria we provide here are useful tools in future analyses of dictionary learning.

In the next section, we give precise statements of our main results, which include an explicit form for $\varepsilon(\delta_1, \delta_2)$. We then prove our main theorem (Thm.~\ref{DeterministicUniquenessTheorem}) in Sec.~\ref{DUT} while stating some additional definitions and lemmas required for the proof, including our main tool from combinatorial matrix analysis (Lem.~\ref{MainLemma}). All other proofs are relegated to the Supplemental. Our approach is a refinement of the one found in \cite{Hillar15} to handle noise and to reduce the sufficient number of samples.
%for all $k < m$ from $N=k{m \choose k}^2$ to $N = m + m(k-1){m \choose k}$. 
Finally, several applications are considered in our discussion, Sec.~\ref{Discussion}.

@@

Before stating our main results, we explain how the spark condition \eqref{SparkCondition} relates to the \emph{lower bound} \cite{Grcar10} of $\mathbf{A}$, which is the largest number $\alpha$ such that \mbox{$\|\mathbf{A}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2$} for all $\mathbf{x} \in \mathbb{R}^m$. By compactness, every injective linear map has a nonzero lower bound; hence, if $\mathbf{A}$ satisfies \eqref{SparkCondition}, then every submatrix formed from $2k$ of its columns or less has a nonzero lower bound. We therefore define the following domain-restricted lower bound of $\mathbf{A}$:
\begin{align*}
L_k(\mathbf{A}) := \frac{1}{\sqrt{k}}\max \{ \alpha : \|\mathbf{A}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2 \text{ for all $k$-sparse } \mathbf{x}\}.
\end{align*} 
Clearly, $L_k(\mathbf{A}) \geq L_{k'}(\mathbf{A})$ whenever $k < k'$, and for any $\mathbf{A}$ satisfying \eqref{SparkCondition}, we have $L_{k'}(\mathbf{A}) > 0$ for all $k' \leq 2k$. Note that subtracting $L_k(\mathbf{A})$ from one yields what is known in the compressed sensing literature as the (asymmetric) lower restricted isometry constant \cite{Blanchard2011, Foucart2009}.

We let $[m]$ denote the set $\{1, \ldots, m\}$ and $2^{[m]}$ ts power set, and we write ${[m] \choose k}$ to denote the set of subsets of $[m]$ of size $k$. Let $\text{Span}\{\mathbf{v}_1, \ldots, \mathbf{v}_\ell\}$ be the $\mathbb{R}$-linear span of vectors $\mathbf{v}_1, \ldots, \mathbf{v}_\ell$. A vector $\mathbf{x}$ is then said to be \emph{supported} on $S \subseteq [m]$ when $\mathbf{x} \in \text{Span} \{\mathbf{e}_j\}_{j\in S}$, where $\mathbf{e}_j$ are the standard $m$-dimensional basis vectors We denote by $\mathbf{M}_j$ the $j$th column of a matrix $\mathbf{M}$ and similarly by $\mathbf{M}_S$ the sub-matrix formed by the columns of $\mathbf{M}$ indexed by $S$. We interpret $\text{Span}\{\mathbf{M}_\emptyset\} = \{\textbf{0}\}$, where $\emptyset$ is the empty set. We say a set system $E \subseteq 2^{[m]}$ is of rank $k$ when $|S| \leq k$ for every $S \in E$ (and $k$-uniform when equality holds). We say $E$ is \emph{regular} when every element of $[m]$ is contained in exactly $\ell$ elements of $E$ for some $\ell > 0$ (for given $\ell$, we say $E$ is $\ell$-regular).

\begin{definition}
Given a set system $E \subseteq 2^{[m]}$, the \emph{star} $F(i)$ centered at $i$ is the set of $S \in E$ containing $i$. We say $E$ satisfies the \emph{singleton intersection property} when $\cap_{S \in F(i)} S = \{i\}$ for all $i \in [m]$. 
\end{definition}

\begin{proposition}
For every $k < m$, there is a regular $k$-uniform set system that satisfies the singleton intersection property.
\end{proposition}
Consider, for instance, the set of consecutive intervals of length $k$ in some cyclic order on $[m]$.

The following result gives an elaborated answer to our question from the introduction. Recall that a set of $k$-sparse vectors is in \emph{general linear position} when any $k$ of them are linearly independent.

[State this without reference to special set $E$]
\begin{theorem}\label{DeterministicUniquenessTheorem}
Fix integers $n$ and $k < m$. Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies \eqref{SparkCondition} and that \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} include at least \mbox{$(k-1){m \choose k}+1$} vectors in general linear position supported on each set in some regular $E \subseteq {[m] \choose k}$ satisfying the singleton intersection property. Then there exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$. %\frac{L_2(\mathbf{A})}{\sqrt{2}}C_1^{-1}$:

Every matrix $\mathbf{B} \in \mathbb{R}^{n \times m}$ for which for some $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{m}$ we have \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$} for all $i$ necessarily satisfies:
\begin{align}\label{Cstable}
\|\mathbf{A}_j-(\mathbf{B}\mathbf{PD})_j\|_2 \leq C_1 \varepsilon \ \ \text{for all $j \in [m]$}
\end{align}
%
for some permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$.\footnote{We note that the condition $\varepsilon < L_2(\mathbf{A})$ above is necessary. When \mbox{$\mathbf{A}$ = $I$} and $\mathbf{x}_i = \mathbf{e}_i$, it turns out that $C_1 = 1$ and there is a $\mathbf{B}$ and $1$-sparse $\mathbf{\hat x}_i$ with $\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$ violating \eqref{Cstable}.}
Furthermore if $\varepsilon < L_{2k}(\mathbf{A}) / C_1$ then $\mathbf{B}$ also satisfies \eqref{SparkCondition} with $L_{2k}(\mathbf{B}\mathbf{PD}) \geq L_{2k}(\mathbf{A}) - C_1 \varepsilon$ and we in fact have:
\begin{align}\label{b-PDa}
\|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i\|_1 &\leq  \left( \frac{ 1+C_1 \|\mathbf{x}_i\|_1 }{ L_{2k}(\mathbf{A}) -  C_1\varepsilon } \right) \varepsilon,
 \ \ \text{for all $i \in [N]$}.
\end{align}
\end{theorem}

@@

An important consequence of Thm.~\ref{DeterministicUniquenessTheorem} is that \eqref{def1} is guaranteed provided $\varepsilon$ does not exceed:
\begin{align}\label{epsdel}
\varepsilon(\delta_1, \delta_2) := \min \left\{ \frac{\delta_1}{ C_1 }, \frac{ \delta_2 L_{2k}(\mathbf{A})}{ 1 + C_1 \left( \max_{i \in [N]} \|\mathbf{x}_i\|_1  + \delta_2 \right) } \right\}.
\end{align}

Note also that Thm~\ref{DeterminsiticUniquenessTheorem} gives conditions for when a single solution to Problem~\ref{InverseProblem} in fact represents \emph{all} possible solutions, related by transformations of the form $PD$.

\begin{corollary}\label{DeterministicUniquenessCorollary}
Given $n, m$, $k < m$ and a regular set system $E \subseteq 2^{[m]}$ of rank $k$ satisfying the singleton intersection property, there are $N =  |E| \left[ (k-1){m \choose k}+1 \right]$ vectors \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} such that every matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying \eqref{SparkCondition} generates a set $Y = \{\mathbf{A}\mathbf{x}_1, \ldots, \mathbf{A}\mathbf{x}_N\}$ with a stable $k$-sparse representation in $m$ dimensions.
\end{corollary}

%\begin{proposition}
%If $E$ is an $\ell$-regular set system of rank $k$ satisfying the singleton intersection property then $|E| \geq \ell m/k$. 
%\end{proposition}
Note that if $E$ is an $\ell$-regular set system of rank $k$ satisfying the singleton intersection property then $|E| \geq \ell m/k$.
This is easy to see by noting that $|E|k \geq \sum_{S \in E}|S| = \sum_{i \in [m]} \deg(i) = \ell m$. In certain cases there exist simple constructions achieving this minimum; for example, when $k = \sqrt{m}$ one can take $E$ to be the rows and columns formed by arranging the elements of $[m]$ in a square grid.

It is straightforward to provide a probabilistic extension of Thm.~\ref{DeterministicUniquenessTheorem} using the following fact in random matrix theory.  The matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfies \eqref{SparkCondition} with probability one (or with ``high probability" for discrete variables) 
provided:
\begin{align}\label{CScondition}
n \geq \gamma k\log\left(\frac{m}{k}\right),
\end{align}
%
where $\gamma$ is a positive constant dependent on the particular distribution from which the entries of $\mathbf{A}$ are sampled i.i.d. (many ensembles suffice, e.g. \cite[Sec.~4]{Baraniuk08}).

In fact, the spark condition can be made explicit.  Let $\mathbf{A}$  be the $n \times m$ matrix of $nm$ indeterminates $A_{ij}$. When real numbers are substituted for all the $A_{ij}$, the resulting matrix satisfies \eqref{SparkCondition} if and only if the following polynomial is nonzero:
\begin{align*}
f(\mathbf{A}) := \prod_{S \in {[m] \choose k}} \sum_{S' \in {[n] \choose k}} (\det \mathbf{A}_{S',S})^2,
\end{align*}
%
where for any $S' \in {[n] \choose k}$ and $S \in {[m] \choose k}$, the symbol $\mathbf{A}_{S',S}$ denotes the submatrix of entries $A_{ij}$ with $(i,j) \in S' \times S$.   We note that the large number of terms in this product is likely necessary due to the NP-hardness of deciding whether $A$ satisfies a given spark condition \cite{tillmann2014computational}.

Since $f$ is a real analytic function, it is enough to show that at least \emph{one} substitution of real numbers satisfies $f(\mathbf{A}) \neq 0$ to conclude that its zeroes form a set with measure zero. Hence, \emph{every} $n \times m$ matrix $\mathbf{A}$ outside such a set satisfies \eqref{SparkCondition} provided \eqref{CScondition} holds for a value of $\gamma$ for \emph{some} distribution. We remark that the precise relationship between $m$, $n$, and $k$ guaranteeing that $f$ is not identically zero is still an open problem in real algebraic geometry [**TODO** cite recent work?]. In any case, we set $\gamma_0$ to be the smallest known such $\gamma$.

It so happens that a similar statement applies to sets of vectors with a stable sparse representation. As in \cite[Sec.~IV]{Hillar15}, consider the ``symbolic" dataset $Y = \{\mathbf{A}\mathbf{x}_1,\ldots,\mathbf{A} \mathbf{x}_N\}$ generated by indeterminate $\mathbf{A}$ and $k$-sparse $\mathbf{x}_1, \ldots, \mathbf{x}_N$.

@@

\begin{proof}[Proof ($k=1$)]
Note that the only 1-uniform set system satisfying the singleton intersection property is $[m]$; hence, we have $\mathbf{x}_i = c_i \mathbf{e}_i$ for $c_i \in \mathbb{R} \setminus \{\mathbf{0}\}$, $i \in [m]$. In this case we may take $C_1$ to be any value satisfying $C_1^{-1} \leq \min_{j \in [m]} |c_{j}|$.

>>>>

Previous theoretical work on Problem \ref{InverseProblem} for the noiseless case $\eta =0$ (e.g., \cite{li2004analysis, Georgiev05, Aharon06, Hillar15}) has shown that a solution (when it exists) is indeed unique provided the $\mathbf{x}_i$ are sufficiently diverse and the generating matrix $\mathbf{A}$ satisfies the \textit{spark condition} from compressed sensing (CS):
\begin{align}\label{SparkCondition}
\mathbf{A}\mathbf{x}_1 = \mathbf{A}\mathbf{x}_2 \implies \mathbf{x}_1 = \mathbf{x}_2, \ \ \ \text{for all $k$-sparse } \mathbf{x}_1, \mathbf{x}_2,
\end{align}
%
which would be, in any case, necessary for uniqueness.

Our main technical finding is that a sufficient diversity of measurements generated from any dictionary satisfying the spark condition 
has a stable sparse representation.  More specifically, such dictionaries are uniquely identifiable from as few as \mbox{$N = m + m(k-1){m \choose k}$} sparse linear combinations of their columns up to an error that is linear in the measurement noise (Thm.~\ref{DeterministicUniquenessTheorem}). In fact, provided $(n, m, k)$ satisfy CS inequality \eqref{CScondition}, in almost all cases the dictionary learning problem is well-posed given enough data (Cor.~\ref{ProbabilisticCor}). Importantly, these guarantees hold without assuming the recovered matrix satisfies the spark condition. Our results  also apply when only an upper bound on the number of dictionary elements is known. The explicit, algorithm-independent criteria we provide should be a useful tool in the theory of sparse dictionary learning.  % (Thm.~\ref{DeterministicUniquenessTheorem2})

In the next section, we give precise statements of our main results, which include an explicit form for $\varepsilon(\delta_1, \delta_2)$. We then prove our main theorem (Thm.~\ref{DeterministicUniquenessTheorem}) in Sec.~\ref{DUT} after stating some additional definitions and lemmas required for the proof, including our main tool from combinatorial matrix analysis (Lem.~\ref{MainLemma}). All other proofs are relegated to the Supplemental. Our approach is a refinement of the one found in \cite{Hillar15} to handle noise and to reduce the sufficient number of samples.
%for all $k < m$ from $N=k{m \choose k}^2$ to $N = m + m(k-1){m \choose k}$. 
Finally, several applications are considered in our discussion, Sec.~\ref{Discussion}.

@@

Before stating our main results, we explain how the spark condition \eqref{SparkCondition} relates to the \emph{lower bound} \cite{Grcar10} of $\mathbf{A}$, which is the largest number $\alpha$ such that \mbox{$\|\mathbf{A}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2$} for all $\mathbf{x} \in \mathbb{R}^m$. By compactness of the unit sphere, every injective linear map has a nonzero lower bound; hence, if $\mathbf{A}$ satisfies \eqref{SparkCondition}, then each submatrix formed from $2k$ of its columns or less has a nonzero lower bound. We therefore define the following domain-restricted lower bound of $\mathbf{A}$:
\begin{align*}
L_k(\mathbf{A}) := \frac{1}{\sqrt{k}}\max \{ \alpha : \|\mathbf{A}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2 \text{ for all $k$-sparse } \mathbf{x}\}.
\end{align*} 
Clearly, $L_k(\mathbf{A}) \geq L_{k'}(\mathbf{A})$ whenever $k < k'$, and for any $\mathbf{A}$ satisfying the spark condition, we have $L_{k'}(\mathbf{A}) > 0$ for all $k' \leq 2k$. Note that subtracting $\sqrt{k} L_k(\mathbf{A})$ [*** TODO: check is this correct ***] from one yields what is known in the compressed sensing literature as the (asymmetric) lower restricted isometry constant \cite{Blanchard2011, Foucart2009}.

An important contribution of our work is to identify combinatorial criteria on the support sets of generating codes that allow for stable sparse 
representations.  Let $[m]$ denote the set $\{1, \ldots, m\}$, $2^{[m]}$ its power set, and ${[m] \choose k}$ the set of subsets of $[m]$ of size $k$. Let $\text{Span}\{\mathbf{v}_1, \ldots, \mathbf{v}_\ell\}$ be the $\mathbb{R}$-linear span of vectors $\mathbf{v}_1, \ldots, \mathbf{v}_\ell$. A vector $\mathbf{x}$ is then said to be \emph{supported} on $S \subseteq [m]$ when $\mathbf{x} \in \text{Span} \{\mathbf{e}_j\}_{j\in S}$, where $\mathbf{e}_j$ are the standard $m$-dimensional basis vectors We denote by $\mathbf{M}_j$ the $j$th column of a matrix $\mathbf{M}$ and similarly by $\mathbf{M}_S$ the sub-matrix formed by the columns of $\mathbf{M}$ indexed by $S$. We interpret $\text{Span}\{\mathbf{M}_\emptyset\} = \{\textbf{0}\}$, where $\emptyset$ is the empty set. We say a hypergraph $E \subseteq 2^{[m]}$ is of rank $k$ when $|S| \leq k$ for every $S \in E$ (and $k$-uniform when equality holds). We say $E$ is \emph{regular} when every element of $[m]$ is contained in exactly $\ell$ elements of $E$ for some $\ell > 0$ (for given $\ell$, we say $E$ is $\ell$-regular).

\begin{definition}
Given a hypergraph $E \subseteq 2^{[m]}$, the \textbf{star} $F(i)$ centered at $i$ is the set of $S \in E$ containing $i$. We say $E$ satisfies the \textbf{singleton intersection property} (SIP) when $\cap_{S \in F(i)} S = \{i\}$ for all $i \in [m]$. 
\end{definition}

\begin{proposition}
For every $k < m$, there is a regular $k$-uniform hypergraph that satisfies the singleton intersection property.
\end{proposition}
Consider, for instance, the set of consecutive intervals of length $k$ in some cyclic order on $[m]$.

Recall that a set of $k$-sparse vectors is in \emph{general linear position} when any $k$ of them are linearly independent.
The following is the technical statement of our main result.

[*** TODO: Combine with Theorem 3 ***]
\begin{theorem}\label{DeterministicUniquenessTheorem}
Fix integers $1 \leq p \leq m$, $n$, and $k < m \leq m' \leq \frac{\ell m - (p-1)}{\ell - 1}$. Suppose $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfies \eqref{SparkCondition} and that \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} include at least \mbox{$(k-1){m \choose k}+1$} vectors in general linear position supported on each set in some $\ell$-regular $E \subseteq {[m] \choose k}$ satisfying the SIP. Then there exists a constant $C_1 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_1$. %\frac{L_2(\mathbf{A})}{\sqrt{2}}C_1^{-1}$:

Every matrix $\mathbf{B} \in \mathbb{R}^{n \times m}$ for which for some $k$-sparse $\mathbf{\hat x}_1, \ldots, \mathbf{\hat x}_N \in \mathbb{R}^{m}$ we have \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$} for all $i$ necessarily satisfies:
\begin{align}\label{Cstable}
\|\mathbf{A}_j-(\mathbf{B}\mathbf{PD})_j\|_2 \leq C_1 \varepsilon \ \ \text{for all $j \in [m]$}
\end{align}
%
for some permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$.
%\footnote{We note that the condition $\varepsilon < L_2(\mathbf{A})$ above is necessary. When \mbox{$\mathbf{A}$ = $I$} and $\mathbf{x}_i = \mathbf{e}_i$, it turns out that $C_1 = 1$ and there is a $\mathbf{B}$ and $1$-sparse $\mathbf{\hat x}_i$ with $\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\hat x}_i\|_2 \leq \varepsilon$ violating \eqref{Cstable}.}
Furthermore if $\varepsilon < L_{2k}(\mathbf{A}) / C_1$ then $\mathbf{B}$ also satisfies \eqref{SparkCondition} with $L_{2k}(\mathbf{B}\mathbf{PD}) \geq L_{2k}(\mathbf{A}) - C_1 \varepsilon$ and we in fact have:
\begin{align}\label{b-PDa}
\|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\hat x}_i\|_1 &\leq  \left( \frac{ 1+C_1 \|\mathbf{x}_i\|_1 }{ L_{2k}(\mathbf{A}) -  C_1\varepsilon } \right) \varepsilon,
 \ \ \text{for all $i \in [N]$}.
\end{align}
\end{theorem}

@@

An important consequence of Thm.~\ref{DeterministicUniquenessTheorem} is that \eqref{def1} is guaranteed provided $\varepsilon$ does not exceed: [*** TODO: verify this is still correct ***]
\begin{align}\label{epsdel}
\varepsilon(\delta_1, \delta_2) := \min \left\{ \frac{\delta_1}{ C_1 }, \frac{ \delta_2 L_{2k}(\mathbf{A})}{ 1 + C_1 \left( \max_{i \in [N]} \|\mathbf{x}_i\|_1  + \delta_2 \right) } \right\}.
\end{align}

% Note also that Thm~\ref{DeterminsiticUniquenessTheorem} gives conditions for when a single solution to Problem~\ref{InverseProblem} in fact represents \emph{all} possible solutions, related by transformations of the form $PD$.

\begin{corollary}\label{DeterministicUniquenessCorollary}
Given $n, m$, $k < m$ and a regular hypergraph $E \subseteq 2^{[m]}$ of rank $k$ satisfying the singleton intersection property, there are $N =  |E| \left[ (k-1){m \choose k}+1 \right]$ vectors \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} such that every matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfying \eqref{SparkCondition} generates a dataset $Y = \{\mathbf{A}\mathbf{x}_1, \ldots, \mathbf{A}\mathbf{x}_N\}$ with a stable $k$-sparse representation in $m$ dimensions.
\end{corollary}

%\begin{proposition}
%If $E$ is an $\ell$-regular hypergraph of rank $k$ satisfying the singleton intersection property then $|E| \geq \ell m/k$. 
%\end{proposition}
Note that if $E$ is an $\ell$-regular hypergraph of rank $k$ satisfying the singleton intersection property then $|E| \geq \ell m/k$.
This is easy to see by noting that $|E|k \geq \sum_{S \in E}|S| = \sum_{i \in [m]} \deg(i) = \ell m$. In certain cases there exist simple constructions achieving this minimum; for example, when $k = \sqrt{m}$ one can take $E$ to be the rows and columns formed by arranging the elements of $[m]$ in a square grid.

It is straightforward to provide a probabilistic extension of Thm.~\ref{DeterministicUniquenessTheorem} using the following fact in random matrix theory.  The matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfies \eqref{SparkCondition} with probability one (or with ``high probability'' for discrete variables) 
provided:
\begin{align}\label{CScondition}
n \geq \gamma k\log\left(\frac{m}{k}\right),
\end{align}
%
where $\gamma$ is a positive constant that depends on the particular distribution from which the entries of $\mathbf{A}$ are sampled i.i.d. (many ensembles suffice, e.g. \cite[Sec.~4]{Baraniuk08}).

In fact, the spark condition can be made explicit.  Let $\mathbf{A}$  be the $n \times m$ matrix of $nm$ indeterminates $A_{ij}$. When real numbers are substituted for $A_{ij}$, the resulting matrix satisfies \eqref{SparkCondition} if and only if the following polynomial is nonzero:
\begin{align*}
f(\mathbf{A}) := \prod_{S \in {[m] \choose k}} \sum_{S' \in {[n] \choose k}} (\det \mathbf{A}_{S',S})^2,
\end{align*}
%
where for any $S' \in {[n] \choose k}$ and $S \in {[m] \choose k}$, the symbol $\mathbf{A}_{S',S}$ denotes the submatrix of entries $A_{ij}$ with $(i,j) \in S' \times S$.   We note that the large number of terms in this product is likely necessary due to the NP-hardness of deciding whether a given matrix $A$ satisfies the spark condition \cite{tillmann2014computational}.

Since $f$ is a real analytic function, it is enough to show that at least one substitution of real numbers satisfies $f(\mathbf{A}) \neq 0$ to conclude that its zeroes form a set with measure zero. Hence, almost every $n \times m$ real matrix $\mathbf{A}$ satisfies \eqref{SparkCondition} provided \eqref{CScondition} holds for a value of $\gamma$ for some distribution. We remark that the precise relationship between $m$, $n$, and $k$ guaranteeing that $f$ is not identically zero is a challenging problem in real algebraic geometry. In any case, we set $\gamma_0$ to be the smallest known such $\gamma$.

It so happens that a similar statement applies to sets of vectors with a stable sparse representation. As in \cite[Sec.~IV]{Hillar15}, consider the ``symbolic'' dataset $Y = \{\mathbf{A}\mathbf{x}_1,\ldots,\mathbf{A} \mathbf{x}_N\}$ generated by indeterminate $\mathbf{A}$ and $k$-sparse $\mathbf{x}_1, \ldots, \mathbf{x}_N$.

@@

\begin{proof}[Proof ($k=1$)]
Note that the only 1-uniform hypergraph satisfying the singleton intersection property is $[m]$; hence, we have $\mathbf{x}_i = c_i \mathbf{e}_i$ for $c_i \in \mathbb{R} \setminus \{\mathbf{0}\}$, $i \in [m]$. In this case we may take $C_1$ to be any value satisfying $C_1^{-1} \leq \min_{j \in [m]} |c_{j}|$.
